Identifying the L1 of non-native writers: the CMU-Haifa system\\[1em] {\large Chris Dyer$^\ast$ Manaal Faruqui$^\ast$ Noam Ordan$^\dagger$ Nathan Schneider$^\ast$\\ Yulia Tsvetkov$^\ast$ Naama Twitto$^\dagger$ Shuly Wintner$^\dagger$}\\[-3em]

### 2013-04-06 paper/paper.tex
from shuly

\nss{shouldn't we use the official bib style file instead of natbib?}

\section{Introduction}
\label{sec:intro}
The task we address in this work is identifying the native language
(\emph{L1}) of non-native English authors. More specifically, given a
dataset
\citep{blanchard-tetreault-higgins-cahill-chodorow:2013:TOEFL11-RR} of
short English essays, composed as part of the TOEFL exam (of English
as a foreign language) by authors whose native language is one out
of~11 possible languages (Arabic, Chinese, French, German, Hindi,
Italian, Japanese, Korean, Spanish, Telugu, and Turkish), our task is
to identify that language.

@@

For classification we use \emph{creg}...

Pre-processing: POS-tagging, etc.

\section{Model Overview}
%\label{sec:features}

\nss{I think we should break this up into sections by feature group (below). 
Here we can talk about learning and regularization and give a high-level overview 
of features.}

We define a large arsenal of features, our motivation being both to
improve the accuracy of classification and to be able to interpret the
characteristics of the language produced by speakers of different
L1s. In this section we define the features and motivate their use.%
\footnote{Whenever counts are mentioned, we use the log of the count
  as the feature.}  
While some of the features were used in the works surveyed in
\Sref{sec:literature}, many are novel, and are inspired by the
features used to identify translationese by \citet{vered:noam:shuly}.
We also report the accuracy of using each feature type, in isolation,
on the training set.

\begin{compactdesc}
\item[Character $n$-grams] The number of character 1-, 2-, and
  3-grams. 69.94\%.
\item[Frequent character $n$-grams] Only those character $n$-grams
  that are observed more than $m$ times in the corpus are
  considered. ??? \nt{this includes 1 to 4 $n$-grams, resulting in 74.12\%}
\item[POS $n$-grams] All essays were tagged with the Stanford part-of-speech tagger \citep{toutanova-03}.
Features were extracted to count every POS 1-, 2-, and 3-gram in each document. 53.92\%.
\item[Document length] The number of tokens in the text. 11.81\%.
\item[Pronouns] The number of each pronoun. 22.81\%.
\item[Punctuation] The number of each punctuation mark. 27.41\%.
\item[Passives] The ratio of verbs to passive verbs. 12.26\%.
\item[Positional token frequency] The choice of the first and last few
  words in a sentence is highly constrained, and may be significantly
  influenced by the authors L1. We use the counts (???)\yt{counts for first two and last three words before the period} of the first
  and last three words in each sentence as features. 53.03\%.
\item[Cohesive markers] These are 40 function words (and short phrases)
  that have a strong discourse function in texts \nt{such as 'however', 'becuase', 'in fact' etc}, contributing to its
  cohesiveness. 25.71\%. \nt{Translators tend to spell out implicit utterances and render them explicitly in the target text (REFERENCE TO BLUM-KULKA); MAYBE APPEND THE LIST IN THE END?} 
\item[Cohesive verbs] \nt{This is a list of manually compiled verbs that serve, like `cohesive markers' to spell out implicit utterances; they include, among others, `indicating', `implying' and `containing'. Same, consider appending the list}. 22.85\%.
\item[Function words] The number of occurrences of each word from a
  pre-defined list of 100 \nt{did you use only 100? it should be 400 and the reference should be to Koppel and Ordan} \yt{we used 100. differences between 100 and >100 were insignificant} most frequent words in English (excluding
  punctuation). 42.47\%.
\item[Contextual function words, bigrams] Pairs consisting of a
  function word from the list mentioned above, along with the POS tag
  of its adjacent word. This feature captures patterns such as verbs
  and the preposition or particle immediately to their right, or nouns
  and the determiner that precedes them. 62.79\%
\item[Contextual function words, trigrams] Same as above, but counting
  \nt{2- or} 3-grams consisting of one or two function words (respectively) and the POS tag of the third word \nt{character?} in the \nt{2- or} 3-gram. 62.32\%.
\item[Lemmas] \nt{these should be actually `stems' as produced by the Stanford tagger} The number of each of the most frequent lemmas in the
  text. ???. 58.95\%.
\item[Prompt] Conjunction of the character $n$-gram features defined
  above with the prompt; since the prompt contributes information on
  the domain, it is likely that some words (and, hence, character
  sequences) will occur more frequently with some prompts than with
  others. 65.09\%.
\item[Misspelling features] ???. 37.29\%.
\item[Brown] ???
\item[Restored] \yt{Counts of substitutions, deletions and insertions of predefined tags that we restored in essays edited with a spelling  corrector and missing all these tags. We define three types of tags: (1) punctuation marks (same list as for punctuation feature), (2) function words (WH-words, prepositions, conjunctions, articles, auxiliary verbs, quantifiers, personal pronouns, possessive pronouns, quantified pronouns), (3) cohesion verbs (lemma, present contiguous/progressive and present simple for each verb). We first remove these tags from texts and then recover the most likely hidden tags in a sequence of words, according to an N-gram language model trained on all essays in the training corpus corrected with a spell checker and containing both words and hidden tags. This feature should capture specific words or punctuation marks that are consistently omitted (deletions), or misused (insertions, substitutions).  To restore hidden tags we use hidden-ngram utility provided in SRILM toolkit\footnote{http://www.speech.sri.com/projects/srilm/manpages/hidden-ngram.1.html} (ref to SRILM). 47.67\%}

\end{compactdesc}

\section{Main Features}\label{sec:mainfeats}

These four feature groups form the core of our model.

\subsection{\feat{POS}: part-of-speech sequences}

\subsection{\feat{FreqChar}: frequent character $n$-grams}

\subsection{\feat{CharPrompt}: character $n$-grams paired with the prompt ID}

\subsection{\feat{Brown}: Brown clusters}

\section{Additional Features}

Each of these adds a small number of parameters to the model. 
We report the empirical improvement that each of these brings 
independently when added to the main features (\Sref{sec:mainfeats}).
The full model combines all features.

\subsection{\feat{CxtFxn}: Contextual function words\nss{does this subsume pronouns?}} \nt{yes}

\subsection{\feat{DocLen}: Document length in tokens}

\subsection{\feat{Misspell}: Spelling correction edits}

\subsection{\feat{Position}: \nss{?}}

\subsection{\feat{Pron}: pronouns\nss{?}} \nt{A list of 25 pronouns in English, again, append list}

\subsection{\feat{PsvRatio}: Ratio of passive to active voice verbs\nss{or is it the proportion of passive verbs?}}

\subsection{\feat{Punct}: Count of each punctuation mark}

\subsection{\feat{Restore}: LM-restored function words}

\subsection{Discarded Features}

\nss{things that didn't help in our preliminary experiments}

\section{Results}
\label{sec:results}

\begin{table}\small\centering
\begin{tabular}{lrcc}
\textbf{Feature Group} & \multicolumn{1}{c}{\textbf{\# Params}} & \textbf{Accuracy (\%)} & \textbf{$\ell_2$} \\
\hline
\feat{POS} & 540,947 & 55.18 & 1.0 \\
+ \feat{FreqChar} & 1,036,871 & 79.55 & 1.0 \\ 
\quad + \feat{CharPrompt} & 2,111,175 & 79.82 & 1.0 \\ 
\qquad + \feat{Brown} & 5,664,461 & 81.09 & 1.0 \\
\end{tabular}
\caption{Dev set accuracy with \textsc{main} feature groups, added cumulatively. 
The number of parameters is always a multiple of 11 (the number of classes). 
Only $\ell_2$ regularization was used for these experiments; 
the penalty was tuned on the dev set as well.}\label{tbl:mainfeats}
\end{table}

\begin{table}\small\centering
\begin{tabular}{lrcc}
\textbf{Feature Group} & \multicolumn{1}{c}{\textbf{\# Params}} & \textbf{Accuracy (\%)} & \textbf{$\ell_2$} \\
\hline
\textsc{main} + \feat{Position} & 6,153,015 & 81.00 & 1.0 \\
\textsc{main} + \feat{PsvRatio} & 5,664,472 & 81.00 & 1.0 \\
\textsc{main} & 5,664,461 & 81.09 & 1.0 \\
\textsc{main} + \feat{DocLen} & 5,664,472 & 81.09 & 1.0 \\
\textsc{main} + \feat{Pron} & 5,664,736 & 81.09 & 1.0 \\
\textsc{main} + \feat{Punct} & 5,664,604 & 81.09 & 1.0 \\
\textsc{main} + \feat{Misspell} & 5,799,860 & 81.27 & 5.0 \\
\textsc{main} + \feat{Restore} & 5,682,589 & 81.36 & 5.0 \\
\textsc{main} + \feat{CxtFxn} & 7,669,684 & 81.73 & 1.0 \\
\end{tabular}
\caption{Dev set accuracy with \textsc{main} features plus additional feature groups, added independently. 
$\ell_2$ regularization was tuned as in \tref{tbl:mainfeats} (two values, 1.0 and 5.0, were tried for each 
configuration; more careful tuning might produce slightly better accuracy).
Results are sorted by accuracy; only three groups exhibited independent improvements over the \textsc{main} feature set.}\label{tbl:addfeats}
\end{table}

\begin{table*}\small\centering
\begin{tabular}{>{\bf}l|r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r|ccc} %from the email sent by the organizers
	& \bf ARA & \bf CHI & \bf FRE & \bf GER & \bf HIN & \bf ITA & \bf JPN & \bf KOR & \bf SPA & \bf TEL & \bf TUR & \bf Precision (\%) & \bf Recall (\%) & \bf $F_1$ (\%) \\
\hline
ARA & 80 & 0 & 2 & 1 & 3 & 4 & 1 & 0 & 4 & 2 & 3 & 80.8 & 80.0 & 80.4 \\
CHI & 3 & 80 & 0 & 1 & 1 & 0 & 6 & 7 & 1 & 0 & 1 & 88.9 & 80.0 & 84.2 \\
FRE & 2 & 2 & 81 & 5 & 1 & 2 & 1 & 0 & 3 & 0 & 3 & 86.2 & 81.0 & 83.5 \\
\hline
GER & 1 & 1 & 1 & 93 & 0 & 0 & 0 & 1 & 1 & 0 & 2 & 87.7 & 93.0 & 90.3 \\
HIN & 2 & 0 & 0 & 1 & 77 & 1 & 0 & 1 & 5 & 9 & 4 & 74.8 & 77.0 & 75.9 \\
ITA & 2 & 0 & 3 & 1 & 1 & 87 & 1 & 0 & 3 & 0 & 2 & 82.1 & 87.0 & 84.5 \\
\hline
JPN & 2 & 1 & 1 & 2 & 0 & 1 & 87 & 5 & 0 & 0 & 1 & 78.4 & 87.0 & 82.5 \\
KOR & 1 & 5 & 2 & 0 & 1 & 0 & 9 & 81 & 1 & 0 & 0 & 80.2 & 81.0 & 80.6 \\
SPA & 2 & 0 & 2 & 0 & 1 & 8 & 2 & 1 & 78 & 1 & 5 & 77.2 & 78.0 & 77.6 \\
\hline
TEL & 0 & 1 & 0 & 0 & 18 & 1 & 2 & 1 & 1 & 73 & 3 & 85.9 & 73.0 & 78.9 \\
TUR & 4 & 0 & 2 & 2 & 0 & 2 & 2 & 4 & 4 & 0 & 80 & 76.9 & 80.0 & 78.4 \\
\end{tabular}
\caption{Official test set confusion matrix with the full model. \nss{which direction is predicted vs. gold?}
Accuracy is 81.5\%.}\label{tbl:matrix}
\end{table*}

@@

\clearpage

>>>>

\section{Introduction}
\label{sec:intro}
The task we address in this work is identifying the native language
(\emph{L1}) of non-native English authors. More specifically, given a
dataset of short English essays
\citep{blanchard-tetreault-higgins-cahill-chodorow:2013:TOEFL11-RR},
composed as part of the \emph{Test of English as a Foreign Language
  (TOEFL)} by authors whose native language is one out of~11 possible
languages (Arabic, Chinese, French, German, Hindi, Italian, Japanese,
Korean, Spanish, Telugu, and Turkish), our task is to identify that
language.

@@

For classification we use \emph{creg}... \nss{Here we can talk about
  learning and regularization and give a high-level overview of
  features.}

All essays were tagged with the Stanford part-of-speech tagger
\citep{toutanova-03}. We did not parse the dataset.

\section{Model Overview}
\label{sec:features}
We define a large arsenal of features, our motivation being both to
improve the accuracy of classification and to be able to interpret the
characteristics of the language produced by speakers of different
L1s.

\subsection{Motivation}
While some of the features were used in the works surveyed in
\Sref{sec:literature}, many are novel, and are inspired by the
features used to identify translationese by \citet{vered:noam:shuly}.
We begin by motivating our choice of features.

\begin{compactdesc}
\item[POS $n$-grams] Part-of-speech $n$-grams were used in various
  text-classification tasks.
\item[Prompt] Since the prompt contributes information on the domain,
  it is likely that some words (and, hence, character sequences) will
  occur more frequently with some prompts than with others. We
  therefore use the prompt ID in conjunction with other features.
\item[Document length] The number of tokens in the text is highly
  correlated with the author's level of fluency, which in turn is
  correlated with the author's L1.
\item[Pronouns] The use of pronoun varies greatly among different
  authors. We use the same list of~25 English pronouns that
  \citet{vered:noam:shuly} use for identifying translationese.
\item[Punctuation] Similarly, different languages use punctuation
  differently, and we expect this to taint the use of punctuation in
  non-native texts.
\item[Passives] English uses passive voice more frequently than other
  languages. Again, the use of passives in L2 can be correlated with
  the author's L1.
\item[Positional token frequency] The choice of the first and last few
  words in a sentence is highly constrained, and may be significantly
  influenced by the author's L1.
\item[Cohesive markers] These are 40 function words (and short
  phrases) that have a strong discourse function in texts (`however',
  `because', `in fact', etc.) Translators tend to spell out implicit
  utterances and render them explicitly in the target text
  \citep{Blum-Kulka:1986}. We use the list
  of~\citet{vered:noam:shuly}.
\item[Cohesive verbs] This is a list of manually compiled verbs that
  are used, like cohesive markers, to spell out implicit utterances
  (`indicate', `imply', `contain', etc.)
\item[Function words] Frequent tokens, which are mostly function
  words, have been used successfully for various text classification
  tasks. \citet{koppel-ordan:2011:ACL-HLT2011} define a list of 400
  such words, of which we only use~100 (using the entire list was not
  significantly different).
\item[Contextual function words] To further capitalize on the ability
  of function words to discriminate, we define pairs consisting of a
  function word from the list mentioned above, along with the POS tag
  of its adjacent word. This feature captures patterns such as verbs
  and the preposition or particle immediately to their right, or nouns
  and the determiner that precedes them. We also define 3-grams
  consisting of one or two function words and the POS tag of the third
  word in the 3-gram.
\item[Lemmas] The content of the text is not considered a good
  indication of the author's L1, but many text categorization tasks
  use lemmas (more precisely, the stems produced by the tagger) as
  features approximating the content.
\item[Misspelling features] Clearly, the spelling errors that learners
  make in English depend on the phonological properties of their
  L1. \sw{???}
\item[Restored tags] We focus on three important token classes defined
  above: punctuation marks, function words and cohesive verbs. We
  first remove words in these classes from the texts, and then recover
  the most likely hidden POS tags in a sequence of words, according to
  an $n$-gram language model trained on all essays in the training
  corpus corrected with a spell checker and containing both words and
  hidden tags. This feature should capture specific words or
  punctuation marks that are consistently omitted (deletions), or
  misused (insertions, substitutions). To restore hidden tags we use
  the \texttt{hidden-ngram} utility provided in SRILM
  \citep{stolcke02srilm}.
\item[Brown clusters] \sw{???}
\end{compactdesc}

\subsection{Main Features}
\label{sec:mainfeats}
First, we use the following four feature types as the core of our
model.  Whenever counts are mentioned, we use the log of the count as
the feature.  We report the accuracy of using each feature type, in
isolation, on the training set.

\begin{compactdesc}
\item[\feat{POS}] Part-of-speech $n$-grams.  Features were extracted
  to count every POS 1-, 2-, and 3-gram in each
  document. 53.92\%. \sw{But the table says 55.18}
\item[\feat{FreqChar}] Frequent character $n$-grams.  We experimented
  with character $n$-grams: The number of character 1-, 2-, and
  3-grams. This yielded 69.94\% accuracy.  We then refined the feature
  to include only those character $n$-grams that are observed more
  than $m$ times in the corpus are considered. \sw{$n$ ranges from~1
    to~4, and $m$ is set to~???. 74.12\%}
\item[\feat{CharPrompt}] Conjunction of the character $n$-gram
  features defined above with the prompt ID. 65.09\%.
\item[\feat{Brown}] Brown clusters. \sw{???}
\end{compactdesc}
\noindent
The accuracy of the classifier on the development set using these four
feature types is reported in \Tref{tbl:mainfeats}.

\begin{table}[hbt]
\small\centering
\begin{tabular}{lrcc}
\textbf{Feature Group} & \multicolumn{1}{c}{\textbf{\# Params}} & \textbf{Accuracy (\%)} & \textbf{$\ell_2$} \\
\hline
\feat{POS} & 540,947 & 55.18 & 1.0 \\
+ \feat{FreqChar} & 1,036,871 & 79.55 & 1.0 \\ 
\quad + \feat{CharPrompt} & 2,111,175 & 79.82 & 1.0 \\ 
\qquad + \feat{Brown} & 5,664,461 & 81.09 & 1.0 \\
\end{tabular}
\caption{Dev set accuracy with \textsc{main} feature groups, added cumulatively. 
  The number of parameters is always a multiple of 11 (the number of classes). 
  Only $\ell_2$ regularization was used for these experiments; 
  the penalty was tuned on the dev set as well.}
\label{tbl:mainfeats}
\end{table}

\subsection{Additional Features}
To the basic set of features we now add more specific,
linguistically-motivated features, each adding a small number of
parameters to the model.  As above, we indicate the accuracy of each
feature type in isolation.

\begin{compactdesc}
\item[\feat{DocLen}] Document length in tokens. 11.81\%.
\item[\feat{Punct}] Counts of each punctuation mark. 27.41\%.
\item[\feat{Pron}] Counts of each pronoun. 22.81\%.
\item[\feat{Position}] Positional token frequency. We use the counts
  for the first two and last three words before the period in each
  sentence as features. 53.03\%.
\item[\feat{PsvRatio}] The proportion of passive verbs out of all
  verbs. 12.26\%.
\item[\feat{CxtFxn}] Contextual function words. Bi-grams yield
  62.79\%, tri-gram 62.32\%.
\item[\feat{Misspell}] Spelling correction edits. \sw{???}. 37.29\%.
\item[\feat{Restore}] Counts of substitutions, deletions and
  insertions of predefined tags that we restored in the texts. 47.67\%
\end{compactdesc}
\noindent
\Tref{tbl:addfeats} reports the empirical improvement that each of
these brings independently when added to the main features
(\Sref{sec:mainfeats}).

\begin{table}[hbt]
\small\centering
\begin{tabular}{lrcc}
\textbf{Feature Group} & \multicolumn{1}{c}{\textbf{\# Params}} & \textbf{Accuracy (\%)} & \textbf{$\ell_2$} \\
\hline
\textsc{main} + \feat{Position} & 6,153,015 & 81.00 & 1.0 \\
\textsc{main} + \feat{PsvRatio} & 5,664,472 & 81.00 & 1.0 \\
\textsc{main} & 5,664,461 & 81.09 & 1.0 \\
\textsc{main} + \feat{DocLen} & 5,664,472 & 81.09 & 1.0 \\
\textsc{main} + \feat{Pron} & 5,664,736 & 81.09 & 1.0 \\
\textsc{main} + \feat{Punct} & 5,664,604 & 81.09 & 1.0 \\
\textsc{main} + \feat{Misspell} & 5,799,860 & 81.27 & 5.0 \\
\textsc{main} + \feat{Restore} & 5,682,589 & 81.36 & 5.0 \\
\textsc{main} + \feat{CxtFxn} & 7,669,684 & 81.73 & 1.0 \\
\end{tabular}
\caption{Dev set accuracy with \textsc{main} features plus additional feature groups, added independently. 
$\ell_2$ regularization was tuned as in \Tref{tbl:mainfeats} (two values, 1.0 and 5.0, were tried for each 
configuration; more careful tuning might produce slightly better accuracy).
Results are sorted by accuracy; only three groups exhibited independent improvements over the \textsc{main} feature set.}
\label{tbl:addfeats}
\end{table}

\subsection{Discarded Features}
We also tried several other feature types that did not improve the
accuracy of the classifier on the development set.
\begin{compactdesc}
\item[Cohesive markers] Counts of each cohesive marker. 25.71\%.
\item[Cohesive verbs] Counts of each cohesive verb. 22.85\%.
\item[Function words] Counts of function words. 42.47\%. This feature
  is subsumed by the highly discriminative \feat{CxtFxn} feature.
\end{compactdesc}

\section{Results}
\label{sec:results}
The full model that we used to classify the test set combines all
features listed in \Tref{tbl:addfeats}. Using all these features, the
accuracy on the development set is \sw{???}, and on the test set it
is~81.5\%. \Tref{tbl:matrix} lists the confusion matrix on the test
set, as well as precision, recall and $F_1$-score for each L1.

\begin{table*}[hbt]
\small\centering
\begin{tabular}{>{\bf}l|r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r@{ }r|ccc} %from the email sent by the organizers
	& \bf ARA & \bf CHI & \bf FRE & \bf GER & \bf HIN & \bf ITA & \bf JPN & \bf KOR & \bf SPA & \bf TEL & \bf TUR & \bf Precision (\%) & \bf Recall (\%) & \bf $F_1$ (\%) \\
\hline
ARA & 80 & 0 & 2 & 1 & 3 & 4 & 1 & 0 & 4 & 2 & 3 & 80.8 & 80.0 & 80.4 \\
CHI & 3 & 80 & 0 & 1 & 1 & 0 & 6 & 7 & 1 & 0 & 1 & 88.9 & 80.0 & 84.2 \\
FRE & 2 & 2 & 81 & 5 & 1 & 2 & 1 & 0 & 3 & 0 & 3 & 86.2 & 81.0 & 83.5 \\
\hline
GER & 1 & 1 & 1 & 93 & 0 & 0 & 0 & 1 & 1 & 0 & 2 & 87.7 & 93.0 & 90.3 \\
HIN & 2 & 0 & 0 & 1 & 77 & 1 & 0 & 1 & 5 & 9 & 4 & 74.8 & 77.0 & 75.9 \\
ITA & 2 & 0 & 3 & 1 & 1 & 87 & 1 & 0 & 3 & 0 & 2 & 82.1 & 87.0 & 84.5 \\
\hline
JPN & 2 & 1 & 1 & 2 & 0 & 1 & 87 & 5 & 0 & 0 & 1 & 78.4 & 87.0 & 82.5 \\
KOR & 1 & 5 & 2 & 0 & 1 & 0 & 9 & 81 & 1 & 0 & 0 & 80.2 & 81.0 & 80.6 \\
SPA & 2 & 0 & 2 & 0 & 1 & 8 & 2 & 1 & 78 & 1 & 5 & 77.2 & 78.0 & 77.6 \\
\hline
TEL & 0 & 1 & 0 & 0 & 18 & 1 & 2 & 1 & 1 & 73 & 3 & 85.9 & 73.0 & 78.9 \\
TUR & 4 & 0 & 2 & 2 & 0 & 2 & 2 & 4 & 4 & 0 & 80 & 76.9 & 80.0 & 78.4 \\
\end{tabular}
\caption{Official test set confusion matrix with the full model. \nss{which direction is predicted vs. gold?}
Accuracy is 81.5\%.}
\label{tbl:matrix}
\end{table*}

\sw{Analysis? Error analysis? Observations?}

@@

%\clearpage
